Phylodynamic inference of contact network parameters through approximate Bayesian computation

### 2016-06-03 doc/thesis/abstract.tex
Updated abstract and a bit more

Models of the spread of disease in a population often make the simplifying
assumption that the population is homogeneously mixed, or is divided into
homogeneously mixed compartments. However, human populations have complex
structures formed by social contacts, which can have a significant influence on
the rate of epidemic spread. Contact network models capture this structure by
explicitly representing each contact that could possibly lead to a
transmission. We developed a method based on approximate Bayesian computation
(ABC) for estimating structural parameters of the contact network underlying an
observed viral phylogeny. The method combines adaptive sequential Monte Carlo
for ABC, Gillespie simulation for propagating epidemics though networks, and a
kernel-based tree similarity score. We used the method to fit the
Barab\'{a}si-Albert network model to simulated transmission trees and applied
it to viral phylogenies estimated from six real-world HIV sequence datasets. On
simulated data, we found that the preferential attachment power and the number
of infected nodes in the network can often be accurately estimated. On the
other hand, the mean degree of the network, as well as the total number of
nodes, appeared to be weakly or non-identifiable with ABC. We observed
substantial heterogeneity in the parameter estimates on real datasets, with
point estimates for the preferential attachment power ranging from 0.06 to
1.05. These results underscore the importance of considering contact structures
when performing phylodynamic inference. Our method offers the potential to
quantitatively investigate the contact network structure underlying viral
epidemics.

>>>>

Models of the spread of disease in a population often make the simplifying
assumption that the population is homogeneously mixed, or is divided into
homogeneously mixed compartments. However, human populations have complex
structures formed by social contacts, which can have a significant influence on
the rate {\color{blue}\uline{and pattern}} of epidemic spread. Contact
{\color{red}\sout{network models}} {\color{blue}\uline{networks}} capture this
structure by explicitly representing each contact that could possibly lead to a
transmission. {\color{blue}\uline{Contact network models parameterize the
structure of these networks, but \ldots }}

We developed a method based on approximate Bayesian computation (ABC) for
estimating structural parameters of the contact network underlying an
observed viral phylogeny. The method combines adaptive sequential Monte Carlo
for ABC, Gillespie simulation for propagating epidemics though networks, and a
{\color{blue}\uline{previously developed}} kernel-based tree similarity score.
{\color{blue}\uline{Our method offers the potential to quantitatively
investigate contact network structure from phylogenies derived from viral
sequence data, as a complementary approach to traditional epidemiological
investigation.}}

We used the method to {\color{red}\sout{fit}}
{\color{blue}\uline{investigate}} the Barab\'{a}si-Albert network model.
{\color{blue}\uline{This model incorporates the mechanism of preferential
attachment observed in real world social and sexual networks, whereby
individuals with a higher number of connections attract new contacts at an
elevated rate (``the rich get richer'').}}
{\color{red}\sout{to simulated transmission trees and applied it to viral
phylogenies estimated from six real-world HIV sequence datasets.}} Using
simulated data, we found that the strength of preferential attachment and the
number of infected nodes in the network could often be accurately estimated. On
the other hand, the mean degree of the network, as well as the total number of
nodes, appeared to be weakly or non-identifiable with ABC.

Finally, the Barab\'{a}si-Albert model was fit to six real-world HIV datasets.
We observed substantial heterogeneity in the parameter estimates on real
datasets, with point estimates for the preferential attachment power ranging
from 0.06 to 1.05. 
{\color{blue}\uline{Point estimates of the strength of preferential attachment 
were higher in injection drug user populations, potentially indicating\ldots}}
{\color{red}\sout{These results underscore the importance of considering
contact structures when performing phylodynamic inference.}}

### 2016-06-03 doc/thesis/discussion.Rnw
A bit more writing
Updated abstract and a bit more

The preferential attachment power \gls{alpha} had a very strong influence on
tree shape in the range of values we considered
(\cref{fig:alphatrees,fig:kpca}). Although the tree kernel was the most
effective classifier for \gls{alpha}, a Sackin's index of tree imbalance
performed nearly as well (\cref{fig:rsquared}). This result was intuitive: high
\gls{alpha} values produce networks with few well-connected ``superspreader''
nodes which are involved in a large number of transmissions, resulting in a
highly unbalanced ladder-like tree structure (\cref{fig:alphatrees}). 
There was no observable bias in the estimates of \gls{alpha} obtained with
\software{netabc}, however the variation in these estimates was higher for
$\alpha < 1$ than for $\alpha \geq 1$ (\cref{fig:abcptm2,tab:abchpd}). The
relationship between \gls{alpha} and the power law exponent \gls{gamma} may
explain this result (\cref{fig:gamma}). The \gls{gamma} values associated with
\gls{alpha} = 0 and \gls{alpha} = 0.5 are nearly identical (about 2.28 for
\gls{alpha} = 0 and 2.33 for \gls{alpha} = 0.5 with \gls{N} = 5000 and \gls{m}
= 2). In other words, the degree distributions of networks with \gls{alpha} $<
1$ are similar to each other, which may result in similarity of corresponding
transmission trees as well. 
%The accuracy of \gls{alpha} estimates also increased when the value of the
%\gls{m} parameter was higher (compare
%\cref{fig:abcptm2,fig:abcptm3,fig:abcptm4}).

@@

>>>>

The preferential attachment power \gls{alpha} had a very strong influence on
tree shape in the range of values we considered
(\cref{fig:alphatrees,fig:kpca}). Although the tree kernel was the most
effective classifier for \gls{alpha}, a Sackin's index of tree imbalance
performed nearly as well (\cref{fig:rsquared}). This result was intuitive: high
\gls{alpha} values produce networks with few well-connected ``superspreader''
nodes which are involved in a large number of transmissions, resulting in a
highly unbalanced ladder-like tree structure (\cref{fig:alphatrees}). 
There was no observable bias in the estimates of \gls{alpha} obtained with
\software{netabc}, however {\color{red}\sout{the variation in these estimates
was higher for $\alpha < 1$ than for $\alpha \geq 1$}} {\color{blue}\uline{there
appeared to be weaker identifiability for \gls{alpha} $<$ 1 than for \gls{alpha}
$\geq$ 1}} (\cref{fig:abcptm2,tab:abchpd}). The relationship between
\gls{alpha} and the power law exponent \gls{gamma} may explain this result
(\cref{fig:gamma}). The \gls{gamma} values associated with \gls{alpha} = 0 and
\gls{alpha} = 0.5 are nearly identical (about 2.28 for \gls{alpha} = 0 and 2.33
for \gls{alpha} = 0.5 with \gls{N} = 5000 and \gls{m} = 2). In other words, the
degree distributions of networks with \gls{alpha} $< 1$ are similar to each
other, which may result in similarity of corresponding transmission trees as
well. 
%The accuracy of \gls{alpha} estimates also increased when the value of the
%\gls{m} parameter was higher (compare
%\cref{fig:abcptm2,fig:abcptm3,fig:abcptm4}).

@@

{\color{blue}\uline{
The posterior mean estimates for the real world \gls{HIV} datasets corresponded
to power law exponents $\gamma$ between 2.09 and 2.49.  TODO
}}

### 2016-06-03 doc/thesis/methods_results.Rnw
Updated abstract and a bit more

\subsubsection*{Joint parameter estimates with kernel-assisted ABC}

@@

{\color{blue}
\uline{Across all simulations, the median [\gls{IQR}] absolute errors of the parameter 
estimates obtained with \software{netabc} were
    \Sexpr{d[,round(median(alpha_error), 2)]} 
    [\Sexpr{d[,round(quantile(alpha_error, 0.25), 2)]} - 
    \Sexpr{d[,round(quantile(alpha_error, 0.75), 2)]}]
for \gls{alpha},
    \Sexpr{d[,round(median(I_error))]} 
    [\Sexpr{d[,round(quantile(I_error, 0.25))]} - 
    \Sexpr{d[,round(quantile(I_error, 0.75))]}]
for \gls{I},
    \Sexpr{d[,round(median(m_error), 2)]} 
    [\Sexpr{d[,round(quantile(m_error, 0.25), 2)]} - 
    \Sexpr{d[,round(quantile(m_error, 0.75), 2)]}]
for \gls{m}, and
    \Sexpr{d[,round(median(N_error))]} 
    [\Sexpr{d[,round(quantile(N_error, 0.25))]} - 
    \Sexpr{d[,round(quantile(N_error, 0.75))]}]
for \gls{N}. These errors comprised, respectively,
    \Sexpr{d[,round(median(alpha_error) * 100 / 2)]}\%,
    \Sexpr{d[,round(median(I_error) * 100 / 4500)]}\%,
    \Sexpr{d[,round(median(m_error) * 100 / 6)]}\%, and
    \Sexpr{d[,round(median(N_error) * 100 / 14500)]}\%
of the regions of nonzero prior density. Average \gls{HPD} interval widths were
    \Sexpr{d[,round(mean(alpha.upper - alpha.lower), 2)]},
    \Sexpr{d[,round(mean(I.upper - I.lower))]},
    \Sexpr{d[,round(mean(m.upper - m.lower), 2)]}, and
    \Sexpr{d[,format(round(mean(N.upper - N.lower)), digits=5)]},
representing 
    \Sexpr{d[,round(mean(alpha.upper - alpha.lower) * 100 / 2)]}\%,
    \Sexpr{d[,round(mean(I.upper - I.lower) * 100 / 4500)]}\%,
    \Sexpr{d[,round(mean(m.upper - m.lower) * 100 / 6)]}\%, and
    \Sexpr{d[,round(mean(N.upper - N.lower) * 100 / 14500)]}\%
of the nonzero prior density regions. Point estimates of \gls{I} were upwardly
biased: \gls{I} was overestimated in
    \Sexpr{d[,sum(I > true_I)]}
out of 
    \Sexpr{nrow(d)}
simulations
    (\Sexpr{round(d[,sum(I < true_I)] * 100 / nrow(d))}\%).
The estimates for \gls{m} and \gls{N} were similar across all simulations
(median [IQR] point estimates
    \Sexpr{d[,round(median(m), 2)]}
    [\Sexpr{d[,round(quantile(m, 0.25), 2)]} - 
    \Sexpr{d[,round(quantile(m, 0.75), 2)]}]
and 
    \Sexpr{d[,round(median(N))]}
    [\Sexpr{d[,round(quantile(N, 0.25))]} - 
    \Sexpr{d[,round(quantile(N, 0.75))]}])
regardless of the true values of any of the \gls{BA} parameters.}

@@

  map1 <- dcast.data.table(d1, dataset~parameter, value.var="mean")
  up1 <- dcast.data.table(d1, dataset~parameter, value.var="hpd.max")
  low1 <- dcast.data.table(d1, dataset~parameter, value.var="hpd.min")
  setkey(map1, dataset)
  setkey(up1, dataset)
  setkey(low1, dataset)

  map2 <- dcast.data.table(d2, dataset~parameter, value.var="mean")
  up2 <- dcast.data.table(d2, dataset~parameter, value.var="hpd.max")
  low2 <- dcast.data.table(d2, dataset~parameter, value.var="hpd.min")
  setkey(map2, dataset)
  setkey(up2, dataset)
  setkey(low2, dataset)
@

We applied kernel-assisted ABC to five published HIV datasets (\cref{tab:data}),
and found substantial heterogeneity among the parameter estimates
(\cref{fig:abchpd,fig:abchpdm2}). Plots of the marginal posterior distributions
for each dataset are shown in
\cref{fig:cuevas,fig:li,fig:niculescu,fig:novitsky,fig:wang}.
Two of the datasets (\textcite{niculescu2015recent, wang2015targeting}) had
estimated $\alpha$ values near unity for the prior allowing $m = 1$ (\gls{MAP}
estimates [95\% \gls{HPD}] 
  \Sexpr{map1["niculescu2015", round(alpha, 2)]} 
  [\Sexpr{low1["niculescu2015", round(alpha, 2)]} - 
   \Sexpr{up1["niculescu2015", round(alpha, 2)]}]
and
  \Sexpr{map1["wang2015", round(alpha, 2)]} 
  [\Sexpr{low1["wang2015", round(alpha, 2)]} -
   \Sexpr{up1["wang2015", round(alpha, 2)]}] respectively).
The MAP estimates did not change appreciably when $m = 1$ was disallowed by the
prior, although the credible interval of the \textcite{niculescu2015recent}
data was narrower
  (\Sexpr{low1["niculescu2015", round(alpha, 2)]} - 
   \Sexpr{up1["niculescu2015", round(alpha, 2)]}).
When $m = 1$ was permitted, the \textcite{li2015hiv, cuevas2009hiv} both had
low estimated $\alpha$ values
  (\Sexpr{map1["li2015", round(alpha, 2)]} 
  [\Sexpr{low1["li2015", round(alpha, 2)]} - 
  \Sexpr{up1["li2015", round(alpha, 2)]}]
and
  \Sexpr{map1["cuevas2009", round(alpha, 2)]} 
  [\Sexpr{low1["cuevas2009", round(alpha, 2)]} -
   \Sexpr{up1["cuevas2009", round(alpha, 2)]}]). 
However, the MAP estimates increased when $m = 1$ was not permitted, although
the HPD intervals remained roughly the same
  (\Sexpr{map2["li2015", round(alpha, 2)]} 
  [\Sexpr{low2["li2015", round(alpha, 2)]} - 
  \Sexpr{up2["li2015", round(alpha, 2)]}]
and
  \Sexpr{map2["cuevas2009", round(alpha, 2)]} 
  [\Sexpr{low2["cuevas2009", round(alpha, 2)]} -
   \Sexpr{up2["cuevas2009", round(alpha, 2)]}]).
The \textcite{novitsky2014impact} data had a fairly low estimated $\alpha$
for both priors on $m$
  (\Sexpr{map1["novitsky2014", round(alpha, 2)]} for $m \geq 1$;
   \Sexpr{map2["novitsky2014", round(alpha, 2)]} for $m \geq 2$).
However, the confidence interval was much wider when $m = 1$ was allowed
  ([\Sexpr{low1["novitsky2014", round(alpha, 2)]} -
    \Sexpr{up1["novitsky2014", round(alpha, 2)]}] for $m \geq 1$ vs.
   [\Sexpr{low2["novitsky2014", round(alpha, 2)]} -
    \Sexpr{up2["novitsky2014", round(alpha, 2)]}] for $m \geq 2$).

For all the datasets except \citeauthor{novitsky2014impact}, estimated values
of $I$ were below 2000 when $m = 1$ was allowed, with relatively narrow HPD
intervals compared to the nonzero prior density region
  (\citeauthor{cuevas2009hiv}, \Sexpr{map1["cuevas2009", floor(I)]} 
  [\Sexpr{low1["cuevas2009", floor(I)]} -
   \Sexpr{up1["cuevas2009", floor(I)]}];
   \citeauthor{niculescu2015recent}, \Sexpr{map1["niculescu2015", floor(I)]}
  [\Sexpr{low1["niculescu2015", floor(I)]} - 
   \Sexpr{up1["niculescu2015", floor(I)]}];
  \citeauthor{li2015hiv}, \Sexpr{map1["li2015", floor(I)]} 
  [\Sexpr{low1["li2015", floor(I)]} -
   \Sexpr{up1["li2015", floor(I)]}];
   \citeauthor{wang2015targeting}, \Sexpr{map1["wang2015", floor(I)]}
  [\Sexpr{low1["wang2015", floor(I)]} - 
   \Sexpr{up1["wang2015", floor(I)]}]).
The \citeauthor{novitsky2014impact} data was the outlier, with a very high
estimated $I$, and HPD interval spanning almost the entire prior region
  (\Sexpr{map1["novitsky2014", floor(I)]} 
  [\Sexpr{low1["novitsky2014", floor(I)]} -
   \Sexpr{up1["novitsky2014", floor(I)]}]).
The $I$ estimates and HPD intervals were generally robust to the choice of
prior on $m$, with slightly narrower HPD intervals (compare
\cref{fig:abchpd,fig:abchpdm2}).

The MAP estimate of $m$ was equal to 1 for all but the
\citeauthor{novitsky2014impact} data, when this value was allowed. However, the
upper bound of the HPD interval was different for each dataset
  (\citeauthor{niculescu2015recent}, \Sexpr{up1["niculescu2015", round(m)]};
   \citeauthor{wang2015targeting}, \Sexpr{up1["wang2015", round(m)]};
   \citeauthor{li2015hiv}, \Sexpr{up1["li2015", round(m)]};
   \citeauthor{cuevas2009hiv}, \Sexpr{up1["cuevas2009", round(m)]}).
When $m = 1$ was disallowed, the MAP for all datasets was either 2 or 3, with
HPD intervals spanning the entire prior region. The estimates for the total
number of nodes $N$ were largely uninformative for all samples, with almost all
MAP estimates greater than 7500 and HPD intervals spanning almost the entire
nonzero prior density region. The only exception was the \citeauthor{li2015hiv}
data, for which the MAP estimate was lower 
  (\Sexpr{map1["li2015", floor(N)]})
when $m = 1$ was allowed.

>>>>

\subsubsection*{Joint parameter estimates with \software{netabc}}

@@

{\color{blue}
\uline{Across all simulations, the median [\gls{IQR}] absolute errors of the parameter 
estimates obtained with \software{netabc} were
    \Sexpr{d[,round(median(alpha_error), 2)]} 
    [\Sexpr{d[,round(quantile(alpha_error, 0.25), 2)]} - 
    \Sexpr{d[,round(quantile(alpha_error, 0.75), 2)]}]
for \gls{alpha},
    \Sexpr{d[,round(median(I_error))]} 
    [\Sexpr{d[,round(quantile(I_error, 0.25))]} - 
    \Sexpr{d[,round(quantile(I_error, 0.75))]}]
for \gls{I},
    \Sexpr{d[,round(median(m_error), 2)]} 
    [\Sexpr{d[,round(quantile(m_error, 0.25), 2)]} - 
    \Sexpr{d[,round(quantile(m_error, 0.75), 2)]}]
for \gls{m}, and
    \Sexpr{d[,round(median(N_error))]} 
    [\Sexpr{d[,round(quantile(N_error, 0.25))]} - 
    \Sexpr{d[,round(quantile(N_error, 0.75))]}]
for \gls{N}. These errors comprised, respectively,
    \Sexpr{d[,round(median(alpha_error) * 100 / 2)]}\%,
    \Sexpr{d[,round(median(I_error) * 100 / 4500)]}\%,
    \Sexpr{d[,round(median(m_error) * 100 / 6)]}\%, and
    \Sexpr{d[,round(median(N_error) * 100 / 14500)]}\%
of the regions of nonzero prior density. Average \gls{HPD} interval widths were
    \Sexpr{d[,round(mean(alpha.upper.95 - alpha.lower.95), 2)]},
    \Sexpr{d[,round(mean(I.upper.95 - I.lower.95))]},
    \Sexpr{d[,round(mean(m.upper.95 - m.lower.95), 2)]}, and
    \Sexpr{d[,format(round(mean(N.upper.95 - N.lower.95)), digits=5)]},
representing 
    \Sexpr{d[,round(mean(alpha.upper.95 - alpha.lower.95) * 100 / 2)]}\%,
    \Sexpr{d[,round(mean(I.upper.95 - I.lower.95) * 100 / 4500)]}\%,
    \Sexpr{d[,round(mean(m.upper.95 - m.lower.95) * 100 / 6)]}\%, and
    \Sexpr{d[,round(mean(N.upper.95 - N.lower.95) * 100 / 14500)]}\%
of the nonzero prior density regions. Point estimates of \gls{I} were upwardly
biased: \gls{I} was overestimated in
    \Sexpr{d[,sum(I > true_I)]}
out of 
    \Sexpr{nrow(d)}
simulations
    (\Sexpr{round(d[,sum(I < true_I)] * 100 / nrow(d))}\%).
The estimates for \gls{m} and \gls{N} were similar across all simulations
(median [IQR] point estimates
    \Sexpr{d[,round(median(m), 2)]}
    [\Sexpr{d[,round(quantile(m, 0.25), 2)]} - 
    \Sexpr{d[,round(quantile(m, 0.75), 2)]}]
and 
    \Sexpr{d[,round(median(N))]}
    [\Sexpr{d[,round(quantile(N, 0.25))]} - 
    \Sexpr{d[,round(quantile(N, 0.75))]}])
regardless of the true values of any of the \gls{BA} parameters.}

@@

  pmean1 <- dcast.data.table(d1, dataset~parameter, value.var="mean")
  up1 <- dcast.data.table(d1, dataset~parameter, value.var="hpd.max.95")
  low1 <- dcast.data.table(d1, dataset~parameter, value.var="hpd.min.95")
  setkey(pmean1, dataset)
  setkey(up1, dataset)
  setkey(low1, dataset)

  pmean2 <- dcast.data.table(d2, dataset~parameter, value.var="mean")
  up2 <- dcast.data.table(d2, dataset~parameter, value.var="hpd.max.95")
  low2 <- dcast.data.table(d2, dataset~parameter, value.var="hpd.min.95")
  setkey(pmean2, dataset)
  setkey(up2, dataset)
  setkey(low2, dataset)
@

We applied kernel-assisted ABC to five published HIV datasets (\cref{tab:data}),
and found substantial heterogeneity among the parameter estimates
(\cref{fig:abchpd,fig:abchpdm2}). Plots of the marginal posterior distributions
for each dataset are shown in
\cref{fig:cuevas,fig:li,fig:niculescu,fig:novitsky,fig:wang}.
Two of the datasets (\textcite{niculescu2015recent, wang2015targeting}) had
estimated $\alpha$ values near unity for the prior allowing $m = 1$ (\gls{MAP}
estimates [95\% \gls{HPD}] 
  \Sexpr{pmean1["niculescu2015", round(alpha, 2)]} 
  [\Sexpr{low1["niculescu2015", round(alpha, 2)]} - 
   \Sexpr{up1["niculescu2015", round(alpha, 2)]}]
and
  \Sexpr{pmean1["wang2015", round(alpha, 2)]} 
  [\Sexpr{low1["wang2015", round(alpha, 2)]} -
   \Sexpr{up1["wang2015", round(alpha, 2)]}] respectively).
The MAP estimates did not change appreciably when $m = 1$ was disallowed by the
prior, although the credible interval of the \textcite{niculescu2015recent}
data was narrower
  (\Sexpr{low1["niculescu2015", round(alpha, 2)]} - 
   \Sexpr{up1["niculescu2015", round(alpha, 2)]}).
When $m = 1$ was permitted, the \textcite{li2015hiv, cuevas2009hiv} both had
low estimated $\alpha$ values
  (\Sexpr{pmean1["li2015", round(alpha, 2)]} 
  [\Sexpr{low1["li2015", round(alpha, 2)]} - 
  \Sexpr{up1["li2015", round(alpha, 2)]}]
and
  \Sexpr{pmean1["cuevas2009", round(alpha, 2)]} 
  [\Sexpr{low1["cuevas2009", round(alpha, 2)]} -
   \Sexpr{up1["cuevas2009", round(alpha, 2)]}]). 
However, the MAP estimates increased when $m = 1$ was not permitted, although
the HPD intervals remained roughly the same
  (\Sexpr{pmean2["li2015", round(alpha, 2)]} 
  [\Sexpr{low2["li2015", round(alpha, 2)]} - 
  \Sexpr{up2["li2015", round(alpha, 2)]}]
and
  \Sexpr{pmean2["cuevas2009", round(alpha, 2)]} 
  [\Sexpr{low2["cuevas2009", round(alpha, 2)]} -
   \Sexpr{up2["cuevas2009", round(alpha, 2)]}]).
The \textcite{novitsky2014impact} data had a fairly low estimated $\alpha$
for both priors on $m$
  (\Sexpr{pmean1["novitsky2014", round(alpha, 2)]} for $m \geq 1$;
   \Sexpr{pmean2["novitsky2014", round(alpha, 2)]} for $m \geq 2$).
However, the confidence interval was much wider when $m = 1$ was allowed
  ([\Sexpr{low1["novitsky2014", round(alpha, 2)]} -
    \Sexpr{up1["novitsky2014", round(alpha, 2)]}] for $m \geq 1$ vs.
   [\Sexpr{low2["novitsky2014", round(alpha, 2)]} -
    \Sexpr{up2["novitsky2014", round(alpha, 2)]}] for $m \geq 2$).

For all the datasets except \citeauthor{novitsky2014impact}, estimated values
of $I$ were below 2000 when $m = 1$ was allowed, with relatively narrow HPD
intervals compared to the nonzero prior density region
  (\citeauthor{cuevas2009hiv}, \Sexpr{pmean1["cuevas2009", floor(I)]} 
  [\Sexpr{low1["cuevas2009", floor(I)]} -
   \Sexpr{up1["cuevas2009", floor(I)]}];
   \citeauthor{niculescu2015recent}, \Sexpr{pmean1["niculescu2015", floor(I)]}
  [\Sexpr{low1["niculescu2015", floor(I)]} - 
   \Sexpr{up1["niculescu2015", floor(I)]}];
  \citeauthor{li2015hiv}, \Sexpr{pmean1["li2015", floor(I)]} 
  [\Sexpr{low1["li2015", floor(I)]} -
   \Sexpr{up1["li2015", floor(I)]}];
   \citeauthor{wang2015targeting}, \Sexpr{pmean1["wang2015", floor(I)]}
  [\Sexpr{low1["wang2015", floor(I)]} - 
   \Sexpr{up1["wang2015", floor(I)]}]).
The \citeauthor{novitsky2014impact} data was the outlier, with a very high
estimated $I$, and HPD interval spanning almost the entire prior region
  (\Sexpr{pmean1["novitsky2014", floor(I)]} 
  [\Sexpr{low1["novitsky2014", floor(I)]} -
   \Sexpr{up1["novitsky2014", floor(I)]}]).
The $I$ estimates and HPD intervals were generally robust to the choice of
prior on $m$, with slightly narrower HPD intervals (compare
\cref{fig:abchpd,fig:abchpdm2}).

The MAP estimate of $m$ was equal to 1 for all but the
\citeauthor{novitsky2014impact} data, when this value was allowed. However, the
upper bound of the HPD interval was different for each dataset
  (\citeauthor{niculescu2015recent}, \Sexpr{up1["niculescu2015", round(m)]};
   \citeauthor{wang2015targeting}, \Sexpr{up1["wang2015", round(m)]};
   \citeauthor{li2015hiv}, \Sexpr{up1["li2015", round(m)]};
   \citeauthor{cuevas2009hiv}, \Sexpr{up1["cuevas2009", round(m)]}).
When $m = 1$ was disallowed, the MAP for all datasets was either 2 or 3, with
HPD intervals spanning the entire prior region. The estimates for the total
number of nodes $N$ were largely uninformative for all samples, with almost all
MAP estimates greater than 7500 and HPD intervals spanning almost the entire
nonzero prior density region. The only exception was the \citeauthor{li2015hiv}
data, for which the MAP estimate was lower 
  (\Sexpr{pmean1["li2015", floor(N)]})
when $m = 1$ was allowed.
